---
title: "Project 3"
author: "Suraj"
date: "24/12/2021"
output:
  pdf_document: default
  html_document: default
---

__Part 1: Data visualization with base R package__

```{r}
cdata <- as.data.frame(CO2)

barplot(table(cdata$Plant),main = "Barplot of Plant")
```
```{r}
barplot(table(cdata$Type),main = "Barplot of Type",col = "lavender",ylim = range(0,50))
```
```{r}
barplot(table(cdata$Treatment),main = "Barplot of Treatment",col = "mistyrose",ylim = range(0,50),las=1)
```

```{r}
hist(cdata$conc,main = "Histogram of Concentration",col = "cornsilk",xlab = "Conc",ylim = range(0,14),las=1)
```
```{r}
hist(cdata$uptake,main = "Histogram of Uptake",col = "lightblue",xlab = "Uptake",xlim = range(0,60))
```
```{r}
plot(cdata$conc,cdata$uptake, main = "Scatterplot of Conc and Uptake",xlab = "Conc",ylab = "Uptake",ylim = range(0,50),las=1)
```
In the scatterplot, we can clearly see that the association between the conc and uptake variable is non linear. And thus, correlation should be calculated using spearman method.

```{r}
cor(cdata$conc,cdata$uptake,method = "spearman")
```
As there was non linear relationship between conc and uptake variable, correlation using spearman method is calculated. The value shows positive, weak relationship between conc and uptake variable.

__Part 2 : Data Visualization with ggplot2 package__

```{r}
library(ggplot2)
ggplot(data = CO2, aes(Plant))+
  geom_bar()+
  labs(title = "Histogram of Plant variable")+
  ylim(0,8)
```
```{r}
ggplot(CO2,aes(Type))+
  geom_bar(fill="lightblue")+
  ggtitle("Histogram of Type variable")+
  ylim(0,50)


```

```{r}
ggplot(CO2,aes(Treatment))+
  geom_bar(fill="steelblue")+
  ggtitle("Histogram of Treatment variable")+
  ylim(0,45)
```

```{r}
ggplot(CO2,aes(conc))+
  geom_histogram(binwidth = 50)+
  ggtitle("Histogram of Conc variable")
```
```{r}
xlab("Conc")
```
```{r}
ggplot(CO2,aes(uptake))+
  geom_histogram(fill="steelblue")+
  ggtitle("Histogram of Treatment variable")+
  xlab("Uptake")+
  ylim(0,8)
```

```{r}
ggplot(CO2,aes(conc,uptake))+
  geom_point()+
  labs(x = "Conc",y = "Uptake")+
  ylim (NA,50)
```
The conc and uptake variable show the non linear association between them. As the non linear association exist, correlation using spearman method is used.

```{r}
cor(cdata$conc,cdata$uptake,method = "spearman")
```
__Part 3: Advantages and Limitations of base R and ggplot2 package__

ggplot2 is an R package for producing statistical or data graphics. It implements a grammar of graphics based on Wilkinson’s book. Comparison of both are as follows:

-The base idea of ggplot2 is to split a chart into graphical objects,data, scale, coordinate system, annotation and think about it separately. On the other hand, function in base R graphics specify everything in the function arguments. In ggplot2 ‘+’ symbol is used to separate the different layers whereas in base R, comma(,) separates the arguments in the function.

-ggplot2 some built in data management. For example ‘bin’ for barplot. It automatically creates the bin and count the each element in the bin. In base R ‘table’ that is outside of the visualization function is used.

-ggplot2 has the consistent syntax. The idea is to create the different layers and ‘+’ symbol separating those symbols. The cons is, it has different type of syntax then rest of R.

-ggplot2 is the powerful function for visualization,can easily create complex graph compared to base R package.

-Base R package is simple, faster and straight forward than ggplot2. ggplot2 is slower than base R package. But the syntax of the Base R package will be heavy for the complex graph.

__Part 4: Text analysis with base/ggplot packages and Social Network Analysis with igraph package__
 
```{r}
library(igraph)
# Loading the termDocMatrix
load("termDocMatrix.rdata")

#Converting to matrix
tmatrix <- as.matrix(termDocMatrix)
class(tmatrix)
```
```{r}
#Getting the term frequency
freq <- sort(rowSums(tmatrix),decreasing = T)
freq
```

```{r}
hist(freq,main = "Histogram of Word Frequency", xlab = "Freq",ylab = NULL,xlim = range(0,100),ylim = range(0,10),las=1)
```
```{r}
hist(freq,main="Histogram of Frequency", xlab = "Words",ylab=NA,xlim = range(0,100),ylim = range(0,10),las = 1)
```
```{r}
#Creating histogram for frequency greater than or equal to 5 using ggplot2 package
dmatrix <- as.data.frame(tmatrix)
freq1 <- rowSums(dmatrix)
freq1
```

```{r}
ggplot(dmatrix[which(freq1 >=5)],aes(freq))+
  geom_histogram()+
  ggtitle("Histogram of Word frequency")+
  labs(x = "Word frequency",y="count")
```

```{r}
library(wordcloud)
```
```{r}
wordcloud(words=names(freq),freq = freq,random.order = T)
```

__Performing Social Network Analysis__

```{r}
#Converting each cell matrix to 1 if greater than 1 and creating a transition matrix
tmatrix[tmatrix > 1] = 1
tranmatrix <- tmatrix %*% t(tmatrix)
tranmatrix[5:9,5:9]
```
```{r}
library(igraph)

```

```{r}
gp <- graph.adjacency(tranmatrix,mode = "undirected",weighted = T)
simplify(gp)
```
```{r}
V(gp)$label <- V(gp)$name
V(gp)$degree <- degree(gp)

set.seed(3952)
layout1 <- layout.fruchterman.reingold(gp)
plot(gp,layout1)
```

Here sna graph is plotted, where nodes are the words in ther termDocMatrix and links defines the relation with other words.


